\textbf{Too Beautiful to be Fake: Attractive Faces are Less Likely to be Judged as Artificially Generated}

### 2022-09-29 submission/manuscript.Rmd
update more

The analysis was carried out using *R 4.2* [@RCoreTeam2022], the *tidyverse* [@wickham2019], and the *easystats* collection of packages [@insightArticle; @bayestestRArticle; @performanceArticle; @parametersArticle; @correlationArticle]. As all the details, scripts and complimentary analyses are open-access, the manuscript will focus on significant results.

@@

Attractiveness had a significant positive and linear relationship ($R^2_{marginal}$ = 2.0%) with the belief that a stimulus was real ($\beta_{poly1} = 16.57$, $95\% CI = [7.33, 25.82]$, $z = 3.51$, $p < .001$) for males, and a quadratic relationship for females ($\beta_{poly2} = 7.82$, $95\% CI = [1.81, 13.84]$, $z = 2.55$, $p = .011$), with both non-attractive and attractive faces being judged as more real. Attractiveness was also found to have a significant positive and quadratic relationship with confidence in judging faces both as real ($\beta_{poly2} = 4.30, $, $95\% CI = [0.97, 7.64]$, $z = 2.53$, $p = .011$) and as fake ($\beta_{poly2} = 5.23, $, $95\% CI = [0.86, 9.60]$, $z = 2.35$, $p = .019$) for females. For males, however, a significant negative and quadratic relationship was found between attractiveness ratings and belief confidence only for faces judged as fake ($\beta_{poly2} = -9.92, $, $95\% CI = [-18.99, -0.86]$, $z = -2.15$, $p = .032$). There was no interaction with reported self-attractiveness.

@@

Trustworthiness, adjusted for beauty and familiarity, had a predominantly significant positive and linear relationship ($R^2_{marginal}$ = 2.0%) with the belief that a stimulus was real ($\beta_{poly1} = 6.44$, $95\% CI = [-0.11, 13.00]$, $z = 1.93$, $p = .0054$) for $p\approx .05$ for females only. No effect on confidence was found for males, whereas a quadratic relationship was found for females for both faces judged as real ($\beta_{poly2} = 6.14$, $95\% CI = [2.13, 10.14]$, $z = 3.00$, $p = .003$) as well as fake ($\beta_{poly2} =6.12$, $95\% CI = [1.49, 10.75]$, $z = 2.59$, $p = .001$), suggesting that non-trustworthy and highly trustworthy faces were rated with more confidence as real or fake than average faces.

@@

Overall social anxiety levels (obtained from aggregating the 2 anxiety scales - *SIAS-6* and *SPS-6*), on the other hand, was significantly associated with the confidence in real judgements only ($\beta_{real} = -0.22$, $95\% CI = [-0.30, -0.14]$, $z = -5.29$, $p < .001$). Similarly, *Intolerance to Uncertainty* was also negatively associated with confidence judgements of faces being real ($\beta_{real} = -0.30$, $95\% CI = [-0.45, -0.15]$, $z = -3.88$, $p < .001$). In particular, confidence in real judgements was found to share predominantly significant positive links with the *Prospective Anxiety* dimension ($\beta_{real} = 1.43$, $95\% CI = [0.10, 2.76]$, $z = 2.10$, $p = .036$), although there was a negative association found with the *Inhibitory Anxiety* dimension ($\beta_{real} = -1.00$, $95\% CI = [-2.01, 0.02]$, $z = -1.93$, $p = 0.054$). There was also a largely significant positive relationship between the *Persecutory Ideation* dimension of *Paranoid Beliefs* with the probability of faces being judged as real ($\beta = 1.87$, $95\% CI = [0.99, 3.54]$, $z = 1.93$, $p = .054$) for $p\approx .05$. However, the negative relationship between confidence judgements and *Paranoid Beliefs* was only found for faces judged as real ($\beta_{real} = -0.21$, $95\% CI = [-0.29, -0.14]$, $z = -5.63$, $p < .001$).

@@

<!-- Attractiveness: females -->
Although attractiveness did not seem to be the primary drive underlying simulation monitoring of face images, we do nonetheless report significant associations, with a different pattern observed depending on the participant's gender. The quadratic relationship found for female participants is aligned with our hypothesis that salient faces (i.e., rated as very attractive or very unattractive) are judged to be more real. The fact that this effect did not reach significance for beauty underlines that attractiveness judgment, and its role in simulation monitoring, is a multidimensional construct that cannot be reduced to physical facial attractiveness, in particular for women [@buunk2002age; @qi2022gender].

@@

<!-- trustworthiness -->
Our results suggested that trustworthiness ratings linearly increased beliefs in realness for females only. Given the evidence suggesting that faces presented as real were rated as more trustworthy [@liefooghe2022faces; @hoogers2021effect; @balas2017], we expected such a link to be present for both genders. One of the underlying mechanisms that possibly contributed to this dimorphism could have been the increased risk-taking aversion reported in females [explained evolutionarily as a compromise to their reproductive potential, @van2020sex], to which perceived facial trustworthiness relates [@hou2019survival]. However, if that was the case, faces judged as highly untrustworthy should have appeared as even more salient (representing an evolutionary threat), and hence be judged as more real. As such, further studies are needed to investigate the causes of the increased simulation monitoring sensitivity to trustworthiness in females.

<!-- familiarity -->
Contrary to our hypothesis, familiarity was not found to be significantly related to simulation monitoring decisions. However, it is to be noted that the distribution of familiarity ratings was strongly skewed, and only a low number of pictures was rated as highly familiar. Future studies should clarify this point by experimentally manipulating familiarity, for instance by modulating the amount of exposure to items before querying the simulation monitoring judgments.

<!-- inter-indiv correlates -->
Regarding the role of inter-individual characteristics in simulation monitoring tendencies, we found that participants with higher scores of honesty-humility, a trait related to an increased risk perception and aversion [@levidi2022understanding; @weller2012honest], displayed a lower confidence in their simulation monitoring judgments. Given the negative correlation between honesty-humility and narcissism [@hodson2018dark], we put this result in perspective with the positive link observed with narcissistic acclaim seeking, confirming the evidence regarding the relationship between narcissistic grandiosity and over-confidence in decision-making [@campbell2004narcissism; @chatterjee2017master; @brunell2017narcissism; @o2021grandiose]. However, an inverse effect was found for other facets of narcissism, such as authoritativeness and thrill seeking. We interpret the effects of the former as related to a higher response assertiveness and, for thrill-seeking, a tolerance of cognitive uncertainty (note that the intolerance to uncertainty scale, which largely measures the anxiety associated with uncertain events, did not show any link). Taken together, these results suggest that participants with low humility, high recognition desires, and high excitement-seeking traits are more confident in their judgment regarding the real or fake nature of ambiguous stimuli. Participants with opposite traits might perceive a higher risk in the decision process and its potential consequences (e.g., being seen as bad at the task at hand), resulting in more conservative confidence ratings.

<!-- AI attitudes -->
Despite the ubiquity of AI, the literature pertaining to the influence of people's AI attitudes on simulation monitoring is scarce. Contrary to our expectations, we did not find evidence for the role of participants' expectations regarding the capabilities of AI technology (in terms of the realism of its productions). Instead, we found that one's enthusiasm about AI technology was related to an increased confidence in simulation monitoring ratings. It is possible that participants with a highly positive attitude towards AI perceive themselves as having greater knowledge about AI and its capabilities [@said2022artificial], hence permitting themselves to be more confident in their simulation monitoring decisions. This result is in line with reports that AI attitudes interacts with people's perceived self-knowledge to influence their perception of the opportunities and risks accorded by AI applications [@@said2022artificial].

<!-- Social phobia -->
However, it is important to note that all of the aforementioned dispositional traits did only influence the confidence judgments, but not the simulation monitoring bias *per se*. In fact, the only trait that modulated the tendency to rate an image as fake or real was the social phobia score, whereby low values were associated with an increased tendency to judge faces as real. Social phobia being characterized as a heightened fear in social interaction contexts [@kolassa2006psychophysiological; @ran2017impact], the nature of the stimuli - pictures of faces - could play a role in this association. However, given the fact that neutral faces are perceived as more threatening by participants with high social phobia [@yoon2008interpreting; @staugaard2010threatening; @peschard2017overestimation], and thus likely more salient, we would have expected high social phobia tendencies to be associated with a bias towards reality judgment. This apparent contradiction could be explained through the lens of fictional reappraisal, an emotion regulation strategy in which the belief that an emotional stimulus is not real is used as a mean to down-regulate the associated emotional reactions [@sperduti2017distinctive; @sperduti2016paradox; @singer2021cognitive]. Moreover, this result is in line with prior findings suggesting that the likelihood of reality beliefs decreased when pictures were emotional, as compared to when they were neutral [@makowski2016phenomenal]. **TODO: can you double check that this last statement is correct?**. Future studies should clarify the role of emotions and attempt at delineating processes related to emotion perception and reaction from emotion regulation.

<!--TO-DO: Uncertainty Intolerance and Paranoid Thinking -->

<!-- Re-exposure Delay -->
On a methodological level, although the order of presentation of the facial images was randomized to reduce effects of adaptation, a shorter re-exposure delay was associated with an increased likelihood of judging the faces as real. This may be because shorter re-exposure delays could have led to the faces being better remembered and appearing more familiar than faces that were displayed after a longer delay, thereby triggering self-referential and autobiographical memory processing during the repeated display [@abraham2009reality; @gobbini2013prioritized; @taylor2009neural]. Although the distribution variance in re-exposure delay is small, recent research using event-related brain potentials report representations of faces to be qualitatively similar regardless of the degree and type of familiarity (e.g., personal or media-based), only distinguishing between familiar and unfamiliar faces [@wiese2021familiarity]. Moreover, fictional stimuli has been shown to up-regulate emotions when associated with familiarity [@sperduti2016paradox; @makowski2017being], thus biasing its salience and perceived realness. Future studies could hence further investigate the modulatory effects of types and degrees of familiarity on perceived realness judgements.

>>>>

The analysis was carried out using *R 4.2* [@RCoreTeam2022], the *tidyverse* [@wickham2019], and the *easystats* collection of packages [@insightArticle; @bayestestRArticle; @performanceArticle; @parametersArticle; @correlationArticle]. As all the details, scripts and complimentary analyses are open-access, the manuscript will focus on significant results. In particular, given the statistical volatility of the findings, which itself could have been a testament to the shifting nature and inter-individuality of our sense of reality, the present study will concentrate on results that provide very strong evidence against the null hypothesis ($p <.01$).

@@

Attractiveness had a significant positive and linear relationship ($R^2_{marginal}$ = 2.0%) with the belief that a stimulus was real ($\beta_{poly1} = 16.57$, $95\% CI = [7.33, 25.82]$, $z = 3.51$, $p < .001$) for males, and a quadratic relationship for females ($\beta_{poly2} = 7.82$, $95\% CI = [1.81, 13.84]$, $z = 2.55$, $p = .011$), with both non-attractive and attractive faces being judged as more real. Attractiveness was also found to have a significant positive and quadratic relationship with confidence in judging faces both as real ($\beta_{poly2} = 4.30, $, $95\% CI = [0.97, 7.64]$, $z = 2.53$, $p = .011$) and as fake ($\beta_{poly2} = 5.23, $, $95\% CI = [0. 86, 9.60]$, $z = 2.35$, $p = .019$) for females. For males, however, a significant negative and quadratic relationship was found between attractiveness ratings and belief confidence only for faces judged as fake ($\beta_{poly2} = -9.92, $, $95\% CI = [-18.99, -0.86]$, $z = -2.15$, $p = .032$). There was no interaction with reported self-attractiveness.

@@

Trustworthiness, adjusted for beauty and familiarity, had a predominantly significant positive and linear relationship ($R^2_{marginal}$ = 2.0%) with the belief that a stimulus was real ($\beta_{poly1} = 6.44$, $95\% CI = [-0.11, 13.00]$, $z = 1.93$, $p = .0054$) for females only. No effect on confidence was found for males, whereas a quadratic relationship was found for females for both faces judged as real ($\beta_{poly2} = 6.14$, $95\% CI = [2.13, 10.14]$, $z = 3.00$, $p = .003$) as well as fake ($\beta_{poly2} =6.12$, $95\% CI = [1.49, 10.75]$, $z = 2.59$, $p = .001$), suggesting that non-trustworthy and highly trustworthy faces were rated with more confidence as real or fake than average faces.

@@

Overall social anxiety levels (obtained from aggregating the 2 anxiety scales - *SIAS-6* and *SPS-6*), on the other hand, was significantly associated with the confidence in real judgements only ($\beta_{real} = -0.22$, $95\% CI = [-0.30, -0.14]$, $z = -5.29$, $p < .001$). Similarly, *Intolerance to Uncertainty* was also negatively associated with confidence judgements of faces being real ($\beta_{real} = -0.30$, $95\% CI = [-0.45, -0.15]$, $z = -3.88$, $p < .001$). In particular, confidence in real judgements was found to share predominantly significant positive links with the *Prospective Anxiety* dimension ($\beta_{real} = 1.43$, $95\% CI = [0.10, 2.76]$, $z = 2.10$, $p = .036$), although there was a negative association found with the *Inhibitory Anxiety* dimension ($\beta_{real} = -1.00$, $95\% CI = [-2.01, 0.02]$, $z = -1.93$, $p = 0.054$). There was also a largely significant positive relationship between the *Persecutory Ideation* dimension of *Paranoid Beliefs* with the probability of faces being judged as real ($\beta = 1.87$, $95\% CI = [0.99, 3.54]$, $z = 1.93$, $p = .054$). However, the negative relationship between confidence judgements and *Paranoid Beliefs* was only found for faces judged as real ($\beta_{real} = -0.21$, $95\% CI = [-0.29, -0.14]$, $z = -5.63$, $p < .001$).

@@

<!-- Attractiveness: females -->
Although attractiveness did not seem to be the primary drive underlying simulation monitoring of face images, we do nonetheless report significant associations, with a different pattern observed depending on the participant's gender. The quadratic relationship found for female participants is aligned with our hypothesis that salient faces (i.e., rated as very attractive or very unattractive) are judged to be more real. The fact that this effect did not reach significance for beauty underlines that attractiveness judgment, and its role in simulation monitoring, is a multidimensional construct that cannot be reduced to physical facial attractiveness, in particular for women [@buunk2002age; @qi2022gender]. This is further evidenced by the observation that female participants were more confident in their judgement of fake faces when they rated the given stimulus very high or low on beauty.

@@

<!-- trustworthiness -->
Our results suggested that trustworthiness ratings linearly increased beliefs in realness for females only (for p-values around .05). Given the evidence positing that faces presented as real were rated as more trustworthy [@liefooghe2022faces; @hoogers2021effect; @balas2017], we expected such a link to be more strongly present for both genders. One of the underlying mechanisms that possibly contributed to this dimorphism could have been the increased risk-taking aversion reported in females [explained evolutionarily as a compromise to their reproductive potential, @van2020sex], to which perceived facial trustworthiness relates [@hou2019survival]. While this hypothesis seems justified, given that female participants' confidence judgements for both fake and real faces were highest when these faces were rated highly trustworthy and untrustworthy, for the hypothesis to hold true, faces judged as highly untrustworthy should have appeared as even more salient (representing an evolutionary threat), and hence be judged as more real. As such, further studies are needed to investigate the causes of the increased simulation monitoring sensitivity to trustworthiness in females.

<!-- familiarity -->
Contrary to our hypothesis, familiarity was not found to be significantly related to simulation monitoring decisions. Interestingly, however, there was a significant relationship between familiarity and confidence judegments only for males, who rated highly familiar faces as real and unfamiliar faces as fake with greater confidence. It is possible that the confidence judgements of females are less malleable, and thus less likely to be influenced by the variability in face familiarity. In fact, this observation is consistent with that of @kaplan2012effect, whose study showed female participants exhibited greater metacognitive awareness in a facial recall task. However, this has not been corroborated by recent research, which found gender differences in the opposite direction instead (males were significantly overconfident regardless of their performance accuracy and the experimental conditions) [@bailey2021gender; @herbst2020gender]. Furthermore, it is to be noted that the distribution of familiarity ratings was strongly skewed, and only a low number of pictures was rated as highly familiar. As such, future studies should clarify this point by experimentally manipulating familiarity, for instance by modulating the amount of exposure to items before querying the simulation monitoring judgments.

<!-- inter-indiv correlates -->
Regarding the role of inter-individual characteristics in simulation monitoring tendencies, we found that participants with higher scores of honesty-humility, a trait related to an increased risk perception and aversion [@levidi2022understanding; @weller2012honest], displayed a lower confidence in their simulation monitoring judgments.

Notably, greater narcissistic tendencies in dimensions such as acclaim seeking and manipulativeness was found to be associated with a higher probability of judging the face stimuli as real. Indeed, this finding is in line with recent research suggesting that narcissists are less likely to engage in cognitive strategies such as reflective thinking [@littrell2020overconfidently; @ahadzadeh2021social] and are more vigilant of their environment [@carolan2017searching; @grapsas2020and; @eddy2021self]. For instance, @de2019facial reports that participants with higher scores in narcissism were faster at accurately recognizing faces with neutral and low-intensity emotional expressions. This hence implies that individuals with greater narcissistic tendencies could have perceived the face images as more salient and engaged in less thoughtful information-processing, thereby leading to their increased perceptions of the displayed stimulus being real.

Moreover, we put the significant positive links between narcissistic acclaim seeking and confidence judgements in perspective with the negative correlation between honesty-humility and narcissism [@hodson2018dark], thus confirming the evidence regarding the relationship between narcissistic grandiosity and over-confidence in decision-making [@campbell2004narcissism; @chatterjee2017master; @brunell2017narcissism; @o2021grandiose]. However, an inverse effect was found for other facets of narcissism, such as authoritativeness. We interpret this correlation as related to a higher response assertiveness. Taken together, these results suggest that participants with low humility and high recognition desires are more confident in their judgment regarding the real or fake nature of ambiguous stimuli. Participants with opposite traits might perceive a higher risk in the decision-making process and its potential consequences (e.g., being seen as bad at the task at hand), resulting in more conservative confidence ratings.

<!-- for thrill-seeking, a tolerance of cognitive uncertainty (note that the intolerance to uncertainty scale, which largely measures the anxiety associated with uncertain events, did not show any link) -->

<!-- AI attitudes -->
Despite the ubiquity of AI, the literature pertaining to the influence of people's AI attitudes on simulation monitoring is scarce. Contrary to our expectations, we did not find evidence for the role of participants' expectations regarding the capabilities of AI technology (in terms of the realism of its productions). Instead, we found that having a generally positive attitude towards AI was related to a decrease in confidence in real judgements. However, it should be noted that our data analyses found the attitudes towards AI scale was better explained as 3 latent dimensions rather than an aggregated score. Indeed, analysing the extracted factors separately revealed that only one's enthusiasm about AI technology was related to an increased confidence in simulation monitoring ratings. This could potentially be because participants with a highly positive attitude towards AI perceive themselves as having greater knowledge about AI and its capabilities [@said2022artificial], hence permitting themselves to be more confident in their simulation monitoring decisions. This result is in line with reports that AI attitudes interacts with people's perceived self-knowledge to influence their perception of the opportunities and risks accorded by AI applications [@@said2022artificial].

<!-- Social Anxiety, Paranoia and Uncertainty Intolerance -->
Unsurprisingly, individuals with more persecutory ideations (a form of paranoia where one believes they will be harmed by others, @buck2019capturing) were more likely to rate the faces presented as real relative to fake. Consistent with previous findings that persecutory delusions are predicted by a greater sense of presence in a VR environment [@freeman2005psychology], it is likely that people with higher levels of paranoia are more sensitive to social threat and thus perceive the face stimuli presented with greater salience and emotionality, hence increasing their beliefs of its realness [@fornells2015people; @freeman2003can]. However, higher levels of persecutory ideation, social anxiety and intolerance of uncertainty were related to lower confidence in real judgements. While this observation appears contradictory, it could be explained through the lens of fictional reappraisal, an emotion regulation strategy in which the belief that an emotional stimulus is not real is used as a mean to down-regulate the associated emotional reactions [@sperduti2017distinctive; @sperduti2016paradox; @singer2021cognitive]. Indeed, previous studies report that individuals with paranoia and anxiety engaged in more affective processing of external ambiguous stimuli [@fornells2015people]. This result is also in line with prior findings suggesting that the likelihood of reality beliefs decreased when pictures were emotional, as compared to when they were neutral [@makowski2016phenomenal]. As such, given the positive correlation between paranoia, social anxiety and uncertainty intolerance [@morrison2014moderating; @counsell2017intolerance; @barnby2020paranoia], it is possible that individuals high in such traits have greater interpersonal sensitivity and perceived the ambiguous face images with greater emotional reactivity, therefore having to engage in emotional regulation strategies that resulted in a lower confidence for real judgements.

<!-- In fact, the only trait that modulated the tendency to rate an image as fake or real was the social phobia score, whereby low values were associated with an increased tendency to judge faces as real. Social phobia being characterized as a heightened fear in social interaction contexts [@kolassa2006psychophysiological; @ran2017impact], the nature of the stimuli - pictures of faces - could play a role in this association. However, given the fact that neutral faces are perceived as more threatening by participants with high social phobia [@yoon2008interpreting; @staugaard2010threatening; @peschard2017overestimation], and thus likely more salient, we would have expected high social phobia tendencies to be associated with a bias towards reality judgment. This apparent contradiction could be explained through the lens of fictional reappraisal,  **TODO: can you double check that this last statement is correct?**. Future studies should clarify the role of emotions and attempt at delineating processes related to emotion perception and reaction from emotion regulation. -->
<!-- However, it is important to note that all of the aforementioned dispositional traits did only influence the confidence judgments, but not the simulation monitoring bias *per se*. -->

<!-- Re-exposure Delay -->
On a methodological level, although the order of presentation of the facial images was randomized to reduce effects of adaptation, participants were more confident in their judgements for faces perceived as real following a shorter re-exposure delay. This may be because shorter re-exposure delays led to the faces being better remembered and appearing more familiar than faces that were displayed after a longer delay, thereby triggering self-referential and autobiographical memory processing during the repeated display [@abraham2009reality; @gobbini2013prioritized; @taylor2009neural]. In fact, this finding is in line with studies that show an up-regulation of emotions when fictional stimuli were associated with familiarity [@sperduti2016paradox; @makowski2017being], thus biasing its salience and perceived realness. However, that being the case, we would have expected shorter re-exposure delays to be related to a higher probability of the faces being rated as real. As such, future studies should further investigate the modulatory effects of types and degrees of familiarity on perceived realness judgements.

<!--a shorter re-exposure delay was associated with an increased likelihood of judging the faces as real.  Although the distribution variance in re-exposure delay is small, recent research using event-related brain potentials report representations of faces to be qualitatively similar regardless of the degree and type of familiarity (e.g., personal or media-based), only distinguishing between familiar and unfamiliar faces [@wiese2021familiarity]. Moreover, fictional stimuli has been shown to up-regulate emotions when associated with familiarity [@sperduti2016paradox; @makowski2017being]. -->
